\bf Hierarchical Network Structure Promotes Dynamical Robustness

### 2015-07-16 tex/overview.tex
update main text

\subsection{Most hierarchical network architecture maximizes robustness}
Examining this expression noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R$.  Given two connected components $C_{\alpha}$ and $C_{\beta}$ with $v_{\alpha}$ and $v_{\beta}$ nodes respectively, we have a maximum of $v_{\alpha} v_{\beta}$ links going from $C_{\alpha}$ to $C_{\beta}$.  Hence, $l \le \sum_{(\alpha,\beta) \in \hier(G)} v_{\alpha} v_{\beta}$.  Since every acyclic digraph can be embedded into a totally ordered
set, we may assume without loss of generality that our components have
been ordered in a way such that, if $(\alpha,\beta) \in \hier(G)$, then $\alpha <
\beta$.  Hence, $l \le l_{max}$ where
$$l_{max} = \sum_{\alpha=1}^{n-1}\sum_{\beta=\alpha+1}^{n}v_{\alpha}
v_{\beta}~=~\frac{1}{2} \left( \sum_{\alpha=1}^{n} v_{\alpha} \right)^2-\frac{1}{2} \sum_{\alpha=1}^{n}
v_{\alpha}^2.$$
Suppose we have a graph $G$ with SCCs $C_1,\ldots,C_n$ and that $G_{\mathrm{tot}}$
is the graph on $n$ nodes with a link from node $i$ to node $j$ whenever $i < j$.  Then we have a graph $G_{\mathrm{max}}$ such that $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$, the components of $G_{\mathrm{max}}$ are also $C_1, \ldots, C_n$, but that includes all possible links between each pair of components. In this case $R(G)~\le~R(G_{\mathrm{max}})$ for all $G \ne G_{\mathrm{max}}$ where $R(G_{\mathrm{max}})$ is equivalent to \ref{eq:robschematic} with $l_{max}$ substituted for $l$
\begin{widetext}
\begin{align} \label{eq:sccmaxrobustness}
R(G_{\mathrm{max}}) =
\frac{\frac{1}{2} ((\sum_{\alpha=1}^n v_{\alpha})^2 - \sum_{\alpha=1}^n v_{\alpha}^2) +
                    \sum_{\alpha=1}^n \connectivity_{\alpha}
                    % \langle R(\hbox{stab}(C_i),\mu') \rangle_1
                R(C_{\alpha},S,\mu_{\alpha},\tau_{\alpha})}
     {\frac{1}{2} ((\sum_{\alpha=1}^n v_{\alpha})^2 - \sum_{\alpha=1}^n v_{\alpha}^2) +
                    \sum_{\alpha=1}^n \connectivity_{\alpha}}.
\end{align}
\end{widetext}
% where $R(G_{\mathrm{max}},S,\mu,\tau) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
%This bound is attained by $G_{\mathrm{max}}$.
% $\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.
\subsection{Simultaneous perturbation of multiple interaction strengths}
This argument also works when relationships between multiple system components are perturbed simultaneously,
although the notation becomes more complicated.  Suppose that we
resample $m$ interactions.  Then the analogue of \ref{eq:singleresamplemutau} is
\begin{equation}\label{eq:multipleresamplemutau}
\begin{aligned}
\tau^{(m)}(A' \mid A) &= \frac{1}{\binom{d}{m}} \sum_{i_1, \ldots, i_m}
  \tau^{(m)}_{i_1, \ldots, i_m} (A' \mid A), \\
\tau^{(m)}_{i_1, \ldots, i_m} (A' \mid A) &=
 \prod_{k=1}^m \rho_{i_k} (a'_{i_k})
 \prod_{j \notin \{i_1, \ldots i_m\}} \delta(a'_j - a_j).
\end{aligned}
\end{equation}
Now define
\begin{widetext}
\begin{equation*}
M = \left\{(m_0, m_1, \ldots, m_n) \,\bigg|\,
m = \sum_{i=0}^n m_i \quad\&\quad
m_0 \le \sum_{i=1}^{n-1} \sum_{j=i+1}^n \ell_{ij} \quad\&\quad
(\forall i \in \{1, \ldots, n\}) \; m_i < \connectivity_i \right\}.
\end{equation*}
\end{widetext}
Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
$m_0$ links between strongly connected components.  Hence, our
weighted average becomes
\begin{widetext}
\begin{equation}\label{eq:robustnessmultiple}
R(G, S, \mu, \tau^{(m)}) =
\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
      {m \choose m_0, m_1, \ldots. m_n}
      \left(m_0 + \sum\limits_{\alpha=1}^n m_\alpha R(C_\alpha, S, \mu, \tau^{(m_\alpha)}) \right)}
     {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
      {m \choose m_0, m_1, \ldots. m_n} m}
\end{equation}
\end{widetext}

@@

This implies that the interaction graphs for systems that are the most robust will maximize the number of links between a given set of SCCs. This analytical result predicts that any network whose associated dynamical system has the interaction graph equivalent to the total ordering will be more robust than those associated to any of the other interaction graphs in \reffigscc{}. The graph associated to the total ordering is the most hierarchical network architecture for any given number of system components like that of \reffigscc{} top for three component systems where the \emph{highest} component in the hierarchy has directed links to all other nodes in the network, the second highest component has directed links to all other nodes in the network except the highest one, \emph{et cetera} (\refsupp{}). Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system. The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.

\subsection{Demonstration of the analytical relationship between hierarchy and robustness}

>>>>

\subsection{Most hierarchical network architecture maximizes dynamical robustness}
Here we show that dynamical robustness is maximized for a given collection of SCCs by maximizing the number of links between them. This is accomplished by demonstrating that any graph $G$ with fewer than the maximum, $l_{\mathrm{max}}$, of links between the given collection of SCCs must have a lower dynamical robustness than the graph, $G_{\mathrm{max}}$, with $l=l_{\mathrm{max}}$ and that this $G_{\mathrm{max}}$ is indeed the graph $G_{\mathrm{tot}}$ which gives the total ordering on the SCCs. Examining the expression in \ref{eq:robschematic} and noting that $R_{\alpha}$ are all strictly less than one proves that networks maximizing $l$, will also maximize $R$.  Given two connected components $C_{\alpha}$ and $C_{\beta}$ with $v_{\alpha}$ and $v_{\beta}$ nodes respectively, we have a maximum of $v_{\alpha} v_{\beta}$ links going from $C_{\alpha}$ to $C_{\beta}$.  Hence, $l \le \sum_{(\alpha,\beta) \in \hier(G)} v_{\alpha} v_{\beta}$.  Since every acyclic digraph can be embedded into a totally ordered
set, we may assume without loss of generality that our components have
been ordered in a way such that, if $(\alpha,\beta) \in \hier(G)$, then $\alpha <
\beta$.  Hence, $l \le l_{\mathrm{max}}$ where
$$l_{\mathrm{max}} = \sum_{\alpha=1}^{n-1}\sum_{\beta=\alpha+1}^{n}v_{\alpha}
v_{\beta}~=~\frac{1}{2} \left( \sum_{\alpha=1}^{n} v_{\alpha} \right)^2-\frac{1}{2} \sum_{\alpha=1}^{n}
v_{\alpha}^2.$$
Suppose we have a graph $G$ with SCCs $C_1,\ldots,C_n$ and that $G_{\mathrm{tot}}$
is the graph on $n$ nodes with a link from node $i$ to node $j$ whenever $i < j$.  Then we have a graph $G_{\mathrm{max}}$ such that $\hier (G_{\mathrm{max}}) = G_{\mathrm{tot}}$, the components of $G_{\mathrm{max}}$ are also $C_1, \ldots, C_n$, but that includes all possible links between each pair of components. In this case $R(G)~\le~R(G_{\mathrm{max}})$ for all $G \ne G_{\mathrm{max}}$ where $R(G_{\mathrm{max}})$ is equivalent to \ref{eq:robschematic} with $l_{\mathrm{max}}$ substituted for $l$
\begin{widetext}
\begin{align} \label{eq:sccmaxrobustness}
R(G_{\mathrm{max}}) =
\frac{\frac{1}{2} ((\sum_{\alpha=1}^n v_{\alpha})^2 - \sum_{\alpha=1}^n v_{\alpha}^2) +
                    \sum_{\alpha=1}^n \connectivity_{\alpha}
                    % \langle R(\hbox{stab}(C_i),\mu') \rangle_1
                R(C_{\alpha},S,\mu_{\alpha},\tau_{\alpha})}
     {\frac{1}{2} ((\sum_{\alpha=1}^n v_{\alpha})^2 - \sum_{\alpha=1}^n v_{\alpha}^2) +
                    \sum_{\alpha=1}^n \connectivity_{\alpha}}.
\end{align}
\end{widetext}
% where $R(G_{\mathrm{max}},S,\mu,\tau) = R_{\mathrm{max}}(C_1, \ldots C_n)$.
%This bound is attained by $G_{\mathrm{max}}$.
% $\langle R(\hbox{stab} (G_{\mathrm{max}}), \mu') \rangle_1 = R_{\mathrm{max}}(C_1, \ldots C_n)$.
\subsection{Simultaneous perturbation of multiple interaction strengths}
This argument also works when relationships between multiple system components are perturbed simultaneously,
although the notation becomes more complicated.  Suppose that we
resample $m$ interactions.  Then the analogue of \ref{eq:singleresamplemutau} is
\begin{equation}\label{eq:multipleresamplemutau}
\begin{aligned}
\tau^{(m)}(A' \mid A) &= \frac{1}{\binom{d}{m}} \sum_{i_1, \ldots, i_m}
  \tau^{(m)}_{i_1, \ldots, i_m} (A' \mid A), \\
\tau^{(m)}_{i_1, \ldots, i_m} (A' \mid A) &=
 \prod_{k=1}^m \rho_{i_k} (a'_{i_k})
 \prod_{j \notin \{i_1, \ldots i_m\}} \delta(a'_j - a_j).
\end{aligned}
\end{equation}
Now define
\begin{widetext}
\begin{equation*}
M = \left\{(m_0, m_1, \ldots, m_n) \,\bigg|\,
m = \sum_{i=0}^n m_i \quad\&\quad
m_0 \le \sum_{i=1}^{n-1} \sum_{j=i+1}^n \ell_{ij} \quad\&\quad
(\forall i \in \{1, \ldots, n\}) \; m_i < \connectivity_i \right\}.
\end{equation*}
\end{widetext}
Then, given $(m_0, m_1, \ldots, m_n) \in M$, there are ${m \choose
m_0, m_1, \ldots. m_n}$ ways of choosing $m_i$ links from $C_i$ and
$m_0$ links between strongly connected components.  Hence, our
weighted average becomes
\begin{widetext}
\begin{equation}\label{eq:robustnessmultiple}
R(G, S, \mu, \tau^{(m)}) =
\frac{\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
      {m \choose m_0, m_1, \ldots. m_n}
      \left(m_0 + \sum\limits_{\alpha=1}^n m_\alpha R(C_\alpha, S, \mu, \tau^{(m_\alpha)}) \right)}
     {\sum\limits_{(m_0, m_1, \ldots m_n) \in M}
      {m \choose m_0, m_1, \ldots. m_n} m}
\end{equation}
\end{widetext}

@@

This analytical result implies that the interaction graphs for systems that are the most robust will maximize the number of links between a given set of SCCs. Because this result is purely topological in nature, it does not depend at all upon any particular details such as the probability distribution from which the component interaction strengths are sampled or the size of the system. Any network whose associated dynamical system has the interaction graph among its SCCs equivalent to the total ordering is more robust than those associated to any of the other interaction graphs in \reffigscc{}. The graph associated to the total ordering is the most hierarchical network architecture for any given number of SCCs like that of \reffigscc{} top for three component systems where the \emph{highest} component in the hierarchy has directed links to all other nodes in the network, the second highest component has directed links to all other nodes in the network except the highest one, \emph{et cetera} (\refsupp{}). The result that dynamical robustness is correlated with network hierarchy therefore applies to an even broader class of dynamical systems than the particular random ensembles we have studied directly.

\subsection{Examples of the analytical relationship between hierarchy and robustness}
